```{r}
here::i_am("output/paper/_si.qmd")
```

- Files for supplementary data must be accompanied by a summary of the file names and types.

# Supplementary information

## Directed Acyclic Graphs

```{r dag-1}
targets::tar_load("all_dags", store = store)
cat(all_dags$chem_to_out)
```

```{r dag-2}
cat(all_dags$chem_to_marker)
```

```{r dag-3}
cat(all_dags$marker_to_out)
```

# Supplementary tables

## Tables for descriptive data

### Study populations

```{r}
targets::tar_load(
  dplyr::contains("codebook_"),
  store = store
)
```

::: {#supptbl-codebook-1}
```{r codebook-1}
codebook_1
```
Codebook for covariates used in $hitrtse \sim edc$
:::

::: {#supptbl-codebook-2}
```{r codebook-2}
codebook_2
```
Codebook for covariates used in $metabolite \sim edc$
:::

::: {#supptbl-codebook-3}
```{r codebook-3}
codebook_3
```
Codebook for covariates used in $hitrtse \sim metabolite$
:::

### Description of chemicals

::: {#supptbl-info-chems}
```{r info-chems}
edcs_info <- myphd::edcs_information() |>
  tidylog::select(-c(chem_id, smiles, ctd_id, exposome_explorer_id)) |>
  tidylog::mutate(
    var = janitor::make_clean_names(
      short_name,
      replace = c(`-` = ""),
      case = "none"
    ) |>
      tolower()
  ) |>
  tidylog::relocate(var, .after = short_name)
gt::gt(edcs_info,
       groupname_col = "class") |>
  gt::sub_values(
    values = c("NA"),
    replacement = ""
  ) |>
  gt::cols_label(
    full_name = "Compound",
    short_name = "Symbol",
    var = "Variable name",
    pubchem_cid = "PubChem CID",
    parental_compound = "Parental compound"
  ) |>
  gt::tab_style(
    locations = gt::cells_row_groups(group = dplyr::everything()),
    style = list(gt::cell_text(weight = "bold"))
  )
```
Information about non-persistent EDCs
:::

### Description of metabolites

::: {#supptbl-info-mets}
```{r info-mets}
mets_info <- readODS::read_ods(
  path = here::here("docs", "steroids.ods")
) |>
  tidylog::select(analyte, acronym, type, hmdb_id, cas_number)
gt::gt(mets_info,
       groupname_col = "type") |>
  gt::sub_values(
    values = c("NA"),
    replacement = ""
  ) |>
  gt::cols_label(
    analyte = "Metabolite",
    acronym = "Symbol",
    hmdb_id = "HMDB ID",
    cas_number = "CAS number"
  ) |>
  gt::tab_style(
    locations = gt::cells_row_groups(group = dplyr::everything()),
    style = list(gt::cell_text(weight = "bold"))
  ) |>
  gt::tab_footnote(
    footnote = "Human Metabolome Database",
    locations = gt::cells_column_labels(
      columns = hmdb_id
    )
  ) |>
  gt::tab_footnote(
    footnote = "Chemical Abstracts Service",
    locations = gt::cells_column_labels(
      columns = cas_number
    )
  )
```
Information about metabolites
:::

::: {#supptbl-lloq-mets}
```{r lloq-mets}
lloq_mets <- load_steroids()$loq |>
  tidylog::mutate(
    metabolite = stringr::str_replace(
      metabolite,
      "^X", ""
    )
  ) |>
  tidylog::rename(
    lloq = "loq"
  )
gt::gt(lloq_mets |> dplyr::arrange(metabolite)) |>
  gt::tab_footnote(
    footnote = "Lower limit of quantification expressed in nanograms per millilitre.",
    locations = gt::cells_column_labels(
      columns = lloq
    )
  )
```
Lower limits of quantification for the glucocorticosteroids
:::

## Tables for main results

### Balancing weights

```{r}
targets::tar_load(
  dplyr::matches("balance_(\\d)$"),
  store = store
)
```

::: {#supptbl-balance-1}
```{r balance-1}
balance_1 |>
  tidylog::group_by(class) |>
  dplyr::arrange(dplyr::desc(Adjusted)) |>
  tidylog::ungroup() |>
  gt::gt(groupname_col = "class") |>
  gt::fmt_number(n_sigfig = 4) |>
  gt::tab_footnote(
    footnote = "N.",
    locations = gt::cells_column_labels(
      columns = c(Unadjusted, Adjusted)
    )
  ) |>
  gt::tab_footnote(
    footnote = "Truncated weights.",
    locations = gt::cells_column_labels(columns = Adjusted)
  )
```
Effective sample size before and after weights estimation for $hitrtse \sim edc$
:::

::: {#supptbl-balance-2}
```{r balance-2}
balance_2 |>
  tidylog::group_by(class) |>
  dplyr::arrange(dplyr::desc(Adjusted)) |>
  tidylog::ungroup() |>
  gt::gt(groupname_col = "class") |>
  gt::fmt_number(n_sigfig = 4) |>
  gt::tab_footnote(
    footnote = "N.",
    locations = gt::cells_column_labels(
      columns = c(Unadjusted, Adjusted)
    )
  ) |>
  gt::tab_footnote(
    footnote = "Truncated weights.",
    locations = gt::cells_column_labels(columns = Adjusted)
  )
```
Effective sample size before and after weights estimation for $metabolite \sim edc$
:::

::: {#supptbl-balance-3}
```{r balance-3}
balance_3 |>
  dplyr::arrange(dplyr::desc(Adjusted)) |>
  gt::gt() |>
  gt::fmt_number(n_sigfig = 4) |>
  gt::tab_footnote(
    footnote = "N.",
    locations = gt::cells_column_labels(
      columns = c(Unadjusted, Adjusted)
    )
  ) |>
  gt::tab_footnote(
    footnote = "Truncated weights.",
    locations = gt::cells_column_labels(columns = Adjusted)
  )
```
Effective sample size before and after weights estimation for $hitrtse \sim metabolite$
:::

```{r}
targets::tar_load(
  dplyr::matches("weights_(\\d)$"),
  store = store
)
```

::: {#supptbl-weights-1}
```{r weights-1}
tidy_res_weighted_fits(
  dat_tbl = weights_1$dat_tbl,
  dat_plt = weights_1$dat_plt,
  sa_var = NULL
)$table
```
Summary statistics of the estimated balancing weights for $hitrtse \sim edc$
:::

::: {#supptbl-weights-2}
```{r weights-2}
tidy_res_weighted_fits(
  dat_tbl = weights_2$dat_tbl,
  dat_plt = weights_2$dat_plt,
  sa_var = NULL
)$table
```
Summary statistics of the estimated balancing weights for $metabolite \sim edc$
:::

::: {#supptbl-weights-3}
```{r weights-3}
tidy_res_weighted_fits(
  dat_tbl = weights_3$dat_tbl,
  dat_plt = weights_3$dat_plt,
  sa_var = NULL
)$table
```
Summary statistics of the estimated balancing weights for $hitrtse \sim metabolite$
:::

## Tables for other results

### Balancing weights for effect modification

```{r}
targets::tar_load(
  dplyr::matches("weights_(\\d)SA$"),
  store = store
)
```

::: {#supptbl-weights-1sa}
```{r weights-1sa}
tidy_res_weighted_fits(
  dat_tbl = weights_1SA$dat_tbl,
  dat_plt = weights_1SA$dat_plt,
  sa_var = "e3_sex"
)$table
```
Summary statistics of the estimated balancing weights for $hitrtse \sim edc \times sex$
:::

::: {#supptbl-weights-2sa}
```{r weights-2sa}
tidy_res_weighted_fits(
  dat_tbl = weights_2SA$dat_tbl,
  dat_plt = weights_2SA$dat_plt,
  sa_var = "e3_sex"
)$table
```
Summary statistics of the estimated balancing weights for $metabolite \sim edc \times sex$
:::

::: {#supptbl-weights-3sa}
```{r weights-3sa}
tidy_res_weighted_fits(
  dat_tbl = weights_3SA$dat_tbl,
  dat_plt = weights_3SA$dat_plt,
  sa_var = "e3_sex"
)$table
```
Summary statistics of the estimated balancing weights for $hitrtse \sim metabolite \times sex$
:::

# Supplementary figures

## Figures for descriptive data

### Description of chemicals

::: {#suppfig-desc-chems}
```{r desc-chems-lod}
#| fig-width: 6
#| fig-height: 7

viz_desc_vars(
  dat = load_dat_request()$dat,
  vars = vars_of_interest(append_to_chem = "cdesc")$chemicals,
  fct_levels = c(1, 2, 3, 4),
  is_chem = TRUE
)$plot
```
Classification of the measured non-persistent EDCs, by cohort.
:::

### Description of metabolites

::: {#suppfig-desc-mets}
```{r desc-mets-lod}
#| fig-width: 6
#| fig-height: 7

viz_desc_vars(
  dat = myphd::extract_cohort(load_steroids()$desc,
                              id_var = id_var,
                              st = 1, en = 3) |>
    tidylog::mutate(
      cohort = dplyr::case_when(
        cohort == "EDE" ~ "EDEN",
        cohort == "KAN" ~ "KANC",
        cohort == "MOB" ~ "MOBA",
        cohort == "RHE" ~ "RHEA",
        .default = cohort
      )
    ),
  vars = vars_of_interest(append_to_chem = NULL)$metabolites |>
    paste0("_cdesc"),
  fct_levels = c(1, 2, 3, 4),
  is_chem = FALSE
)$plot
```
Classification of the measured glucocorticosteroids, by cohort.
:::

## Figures for other results

### Marginal contrasts for effect modification

```{r}
targets::tar_load(
  dplyr::matches("marginal_comparisons_(\\d)SA$"),
  store = store
)
```

::: {#suppfig-marginal-1sa}
```{r marginal-1sa}
#| fig-width: 6
#| fig-height: 7

tidy_res_meffects(
  df = marginal_comparisons_1SA,
  sa_var = "e3_sex",
  outcome = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```
Marginal contrast results for $hitrtse \sim edc \times sex$
:::

::: {#suppfig-marginal-2sa}
```{r marginal-2sa}
#| fig-width: 9
#| fig-height: 12

tidy_res_meffects(
  df = marginal_comparisons_2SA |>
    tidylog::filter(
      !grepl("metabolism", outcome)
    ),
  sa_var = "e3_sex",
  outcome = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```
Marginal contrast results for $metabolite \sim edc \times sex$
:::

::: {#suppfig-marginal-3sa}
```{r marginal-3sa}
#| fig-width: 6
#| fig-height: 5

tidy_res_meffects(
  df = marginal_comparisons_3SA,
  sa_var = "e3_sex",
  outcome = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```
Marginal contrast results for $hitrtse \sim metabolite \times sex$
:::

### Sensitivity to exposure assessment
