---
title: "Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and attentional function: A study based on the parametric g-formula"

# metadata-files:
#   - ../../docs/authors.yml

format:
  html: default
  docx:
    # reference-doc: reference_document.docx
    page-width: 8.3
    citeproc: true
    mermaid:
      theme: neutral
  pdf:
    documentclass: article
    geometry:
      - top=30mm
      - left=20mm
      - heightrounded
    pdf-engine: lualatex
    number-sections: true
    toc: false
    keep-tex: false
    include-in-header:
      text: |
        \usepackage{lineno}
        \usepackage{rotating}
        \usepackage{float}
        \usepackage{typearea}
    include-before-body:
      text: |
        \linenumbers

crossref:
  custom: 
    - kind: float
      key: supptbl
      latex-env: supptbl
      reference-prefix: Table S
      space-before-numbering: false
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Figure S
      space-before-numbering: false

bibliography: bibliography.bib
csl: american-journal-of-epidemiology.csl

link-citations: true
linkcolor: blue

filters:
  - acronyms
  - ../../../../all_phd/_extensions/wordcount/wordcount.lua

acronyms:
  fromfile: ../../../../all_phd/acronyms.yml
  insert_links: false
  insert_loa: false
  sorting: alphabetical

editor_options: 
  chunk_output_type: console

execute:
  echo: false
  warning: false
  error: false
  message: false
---

```{r}
here::i_am("output/paper/manuscript.qmd")
```

```{r vars-methods}
source(here::here("R/dictionaries.R"))
source(here::here("R/data.R"))
source(here::here("R/utils.R"))
source(here::here("R/proc_res.R"))

is_hpc <- FALSE
store <- ifelse(
  is_hpc, 
  paste0(
    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
    "lorenzoF/data/data_paper3/_targets/_targets_res"
  ), 
  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
)
id_var <- "HelixID"
num_digits_est = 3
num_digits_sig = 2
params_ <- params(is_hpc = is_hpc)
threshold_within_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_within
threshold_overall_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_overall
threshold_within_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_within
threshold_overall_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_overall
knn_k <- params_$steps$rq2$preproc_outcome$missings$k
```

```{r}
format_res <- function(df, exposure, outcome) {
  # df: variable, val, estimate, outcome, class
  # exposure in variable, outcome in outcome
  
  val <- df[df$variable == exposure & df$outcome == outcome, ]$val
  val <- stringr::str_replace(
    string = val,
    pattern = "\\*",
    replacement = ""
  )
  val1 <- stringr::str_split(val, " ", n = 2)[[1]][1]
  val2 <- stringr::str_split(val, " ", n = 2)[[1]][2]
  glue::glue(
    "MC: {val1} and $95\\%$ \\acr{{ci}}: {val2}"
  )
}
```

```{r}
targets::tar_load(
  dplyr::matches("marginal_comparisons_(\\d)$"),
  store = store
)
meffects1 <- tidy_res_meffects(
  df = marginal_comparisons_1,
  sa_var = NULL,
  outcome = c("hitrtse"),
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$numerical_results
meffects2 <- tidy_res_meffects(
  df = marginal_comparisons_2,
  sa_var = NULL,
  outcome = c(
    "cortisol production",
    "cortisone production",
    "corticosterone production"
  ),
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$numerical_results
meffects3 <- tidy_res_meffects(
  df = marginal_comparisons_3,
  sa_var = NULL,
  outcome = c("hitrtse"),
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$numerical_results
```

# Abstract {.unnumbered}

Evidence suggests that {{< acr edcs >}} may perturb the {{< acr hpa >}} axis, which has a major role in brain development. We aimed to evaluate the effects of childhood exposure to organophosphate pesticides, phenols, and phthalate metabolites, on urinary glucocorticosteroids and inattention in children using data from the {{< acr helix >}} cohort. We used the parametric g-formula to estimate effects between {{< acr edcs >}}, glucocorticosteroids, and {{< acr hitrtse >}}, a measure of inattention from the {{< acr ant >}}, and tested for possible effect modification by sex. We observed a positive {{< acr mc >}} for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`), on {{< acr hitrtse >}}, indicating lower attention. Several {{< acr edcs >}} were also associated with a positive {{< acr mc >}} for cortisone, cortisol, and corticosterone production. Increased levels of the glucocorticosteroids had no effect on {{< acr hitrtse >}}, although we found a possible effect modification by sex. Our results suggest that multiple {{< acr edcs >}} might interfere with inattention in children and with the homeostasis of the {{< acr hpa >}} axis.

{{< pagebreak >}}

The prevalence of several neurodevelopmental disorders has increased in the pediatric population [@GrandjeanLandrigan:2014], and multiple environmental pollutants may play a role in the increased rates of these disorders [@RamirezGalvez-OntiverosGonzalez-Domenech:2022]. Multiple {{< acr edcs first_use=true >}}, ubiquitous chemicals present in many every-day products and diet, are capable of interfering with the endocrine system, and have shown associations with childhood neurodevelopment and behavior [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @TewarAuingerBraun:2016; @YuDuChiou:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @FurlongHerringBuckley:2017; @Braun:2017; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019; @ShoaffCoullWeuve:2020; @OhKimKannan:2023; @VilmandBeckBilenberg:2023]. Although both pregnancy and early infancy are crucial stages of (neuro)development, most of the available literature is focused on the effects of prenatal exposure to {{< acr edcs >}} on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].

One group of {{< acr edcs >}} that may have a deleterious effect on neurodevelopment is the {{< acr opps >}}, although the few studies assessing exposure during childhood and through the use of biomarkers suffered from a series of limitations, including a small sample size [@RamirezGalvez-OntiverosGonzalez-Domenech:2022]. Exposure to phthalates and their metabolites during childhood and early adolescence has also been associated with several adverse neurodevelopmental outcomes, but these studies were limited to few phthalate metabolites and small study populations [@RamirezGalvez-OntiverosGonzalez-Domenech:2022]. The effects of exposure to {{< acr bpa >}} during childhood on cognitive functions are still unclear [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].

Moreover, little is known about the biological mechanisms of action [@RamirezGalvez-OntiverosGonzalez-Domenech:2022]. There is some toxicological evidence, however, that exposure to certain {{< acr edcs >}}, specifically phthalates, might interfere with the {{< acr hpa first_use=true >}} axis and might interact with the glucocorticoid receptor [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023]. The HPA axis, which can be activated by stress, is responsible for the production of glucocorticosteroids. The brain, and its proper functioning, is a potential target, due to the presence of receptors for these hormones [@LupienMcEwenGunnar:2009; @SunLiJin:2018]. Glucocorticosteroids are necessary for brain maturation, although their under- or over-production might interfere with its normal development and ultimately lead to long-term impaired functioning [@LupienMcEwenGunnar:2009; @SearsLiuLanphear:2023].

Taken together, these results suggest that the negative influence of exposure to certain {{< acr edcs >}} on neurodevelopmental outcomes might be mediated, at least partially, by disruption of the {{< acr hpa >}} axis' homeostasis.
In the present study, we thus estimated cross-sectional associations between 1) non-persistent {{< acr edcs >}} and attentional function, 2) non-persistent {{< acr edcs >}} and glucocorticosteroids, and 3) glucocorticosteroids and attentional function, using the parametric g-formula and {{< acr mcs first_use=true >}}, in children of a large network of cohorts in Europe.

# Methods {#sec-methods}

## Study population and design {#sec-design}

The {{< acr helix first_use=true >}} project aims to characterize early-life exposures and their potential association with endogenous biomarkers and health outcomes [@VrijheidSlamaRobinson:2014]. It consists of six existing population-based birth cohort studies across Europe: BiB (Born in Bradford, UK) [@WrightSmallRaynor:2013], EDEN (Study of determinants of pre- and postnatal developmental, France) [@HeudeForhanSlama:2016], INMA (Environment and Childhood, Spain) [@GuxensBallesterEspada:2012], KANC (Kaunas Cohort, Lithuania) [@GrazulevicieneDanileviciuteNadisauskiene:2009], MoBa (The Norwegian Mother and Child Cohort Study, Norway) [@MagnusIrgensHaug:2006], and Rhea (Mother–Child Cohort in Crete, Greece) [@ChatziPlanaDaraki:2009]. The {{< acr helix >}} subcohort of 1,301 mother-child pairs was fully characterized for the external and internal exposome, including exposure and omics biomarkers during childhood [@MaitreBontCasas:2018]. Eligibility criteria for inclusion in the {{< acr helix >}} subcohort included: a) age 6-11 years, with a preference for 7-9 years; b) availability of sufficient stored pregnancy blood and urine samples; c) availability of complete address history from first to last follow-up; d) no serious health problems, which might affect the results of the clinical testing. Ethical permission was obtained from the relevant authorities in the corresponding country.

## Variables {#sec-vars}

### Endocrine disrupting chemicals {#sec-edcs}

```{r}
info_edcs <- myphd::edcs_information()
print_tidy_abbreviations_edcs <- function(df, class_) {
  tmp <- df |>
    dplyr::filter(class == class_) |>
    dplyr::select(short_name, full_name) |>
    dplyr::arrange(full_name) |>
    dplyr::mutate(
      ret = glue::glue("{full_name} ({short_name})")
    )
  tmp$ret |>
    paste0(collapse = ", ")
}
```

Children were assessed between December 2013 and February 2016, and assessments included neurological testing and urine collection. Urine samples of the night before and the first morning void on the day of the visit were combined to provide a more reliable exposure assessment. Non-persistent {{< acr edcs >}} assessed in the urine samples included phthalate metabolites, phenols, and {{< acr op >}} pesticide metabolites. A list of the environmental chemicals determined in urine samples and used for the present study is given in @supptbl-info-chems. Briefly, we analyzed a total of 7 phenols (`r print_tidy_abbreviations_edcs(info_edcs, "Phenols")`), 6 non-specific organophosphate pesticide metabolites (`r print_tidy_abbreviations_edcs(info_edcs, "OP pesticide metabolites")`), and 10 phthalate metabolites (`r print_tidy_abbreviations_edcs(info_edcs, "Phthalate metabolites")`) originating from 6 distinct phthalate parent compounds.
The laboratory protocols for the analysis are described elsewhere [@HaugSakhiCequier:2018].

### Glucocorticosteroids {#sec-steroids}

```{r}
info_steroids <- readODS::read_ods("../../docs/steroids.ods")
print_tidy_abbreviations_steroids <- function(df, score) {
  cortisol_prod <- c(
    "20aDHF", "20bDHF", "5bDHF", "5aTHF", "5bTHF",
    "6OHF",
    "5a20acortol", "5a20bcortol", "5b20acortol", "5b20bcortol"
  )
  cortisone_prod <- c(
    "20aDHE", "20bDHE", "5aTHE", "5bTHE",
    "6OHE",
    "5b20acortolone", "5b20bcortolone"
  )
  corticosterone_prod <- c(
    "5aTHB", "5bTHB", "A", "17-DO-cortolone"
  )
  tmp <- df |>
    dplyr::filter(dplyr::case_when(
      score == "cortisol_prod" ~ acronym %in% cortisol_prod,
      score == "cortisone_prod" ~ acronym %in% cortisone_prod,
      score == "corticosterone_prod" ~ acronym %in% corticosterone_prod,
    )) |>
    dplyr::select(analyte, acronym) |>
    dplyr::arrange(analyte) |>
    dplyr::mutate(
      ret = glue::glue("{analyte} ({acronym})")
    )
  tmp$ret |>
    paste0(collapse = ", ")
}
```

Urine samples of the night before the day of the visit were used to measure levels of the glucocorticosteroids. These included glucocorticosteroids, glucocorticosteroid metabolites, glucocorticosteroid precursors, glucocorticosteroid precursor metabolites, androgens, and androgen metabolites. A list of the glucocorticosteroids determined in urine samples and used for the present study is given in @supptbl-info-mets.

To assess the levels of glucocorticosteroids and their metabolites, LC-MS/MS analysis was applied at the Applied Metabolomics Research Group, IMIM (Hospital del Mar Medical Research Institute). The laboratory protocols for the analysis are described elsewhere [@MarcosRenauCasals:2014; @Gomez-GomezPozo:2020].

Three additional markers, total cortisol production, total cortisone production, and total corticosterone production, were computed based on the following: cortisol production as the sum of cortisol and its metabolites (`r print_tidy_abbreviations_steroids(info_steroids, "cortisol_prod")`), cortisone production as the sum of cortisone and its metabolites (`r print_tidy_abbreviations_steroids(info_steroids, "cortisone_prod")`), and corticosterone production as the sum of `r print_tidy_abbreviations_steroids(info_steroids, "corticosterone_prod")`.

### Attentionall function {#sec-neurodevelopment}

Cognitive and motor function outcomes were assessed with standardized, non-linguistic, and culturally blind computer tests, including the {{< acr ant first_use=true >}} [@RuedaFanMcCandliss:2004], which provides a measure of efficiency of attentional function. The tests were administered in a standardized way, and with minimal interference from the field workers. Further information can be found in @MaitreBontCasas:2018. The outcome of interest for the present study is the {{< acr hitrtse first_use=true >}} [@SunyerEsnaolaAlvarez-Pedrerol:2015], a measure of response speed consistency throughout the test. A high {{< acr hitrtse >}} indicates highly variable reaction times, and is considered a measure of inattentiveness.

### Confounders {#sec-confounders}

For each research question, defined by a specific type of exposure and outcome, the minimal set of covariates for inclusion in the analyses was selected on the basis of a {{< acr dag >}} built with `DAGitty` [@TextorvanderZanderGilthorpe:2016] and `ggdag` [@Barrett:2023]. The sets of covariates were selected to estimate the total effect of the exposure on the outcome. For effect estimation of the {{< acr edcs >}} on glucocorticosteroids and of glucocorticosteroids on {{< acr hitrtse >}}, these sets were also sufficient to estimate direct effects. Sample-specific creatinine values were used to adjust for possible dilution effects. Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure. Common confounders were cohort, ethnicity, sex, age, height, weight, and head circumference of the child, consumption of fish, fruit, vegetables, organic food, anf fast food, maternal tobacco consumption, family financial situation and affluence scale (FAS). Models for estimating the effects of {{< acr edcs >}} on {{< acr hitrtse >}} were further adjusted for child breastfeeding, prenatal maternal active and passive smoking, urine creatinine, child mood and rest before assessment, child neuropsychological diagnosis, marital status, season, and fasting time before assessment. Models for estimating the effects of {{< acr edcs >}} on glucocorticosteroids were further adjusted for urine creatinine, season, and fasting time before assessment. Models for estimating the effects of glucocorticosteroids on {{< acr hitrtse >}} were further adjusted for child breastfeeding, prenatal maternal active and passive smoking, marital status, {{< acr edcs >}}, urine creatinine, child mood and rest before assessment, and child neuropsychological diagnosis. The adjustment sets are provided in the Supplementary Material as text files compatible with DAGitty. Codebooks for the used covariates, by research question, are provided in Supplementary Tables [-@supptbl-codebook-1; -@supptbl-codebook-2; -@supptbl-codebook-3].

## Statistical methods {#sec-stat-methods}

### Data pre-processing {#sec-dat-preproc}

Concentrations of the glucocorticosteroids were classified as quantifiable, below the {{< acr loq >}}, possible interference or out of range, and not detected. For each metabolite, we computed the fraction of values below the {{< acr loq >}} and not detected, both within each cohort and overall. We proceeded to impute these values using half the value of the corresponding {{< acr loq >}}, for those metabolites that had less than `r threshold_within_lod`% of missings within each cohort and `r threshold_overall_lod`% of missings overall. Information about the {{< acr lloq >}} for the glucocorticosteroids is provided in @supptbl-lloq-mets. The remaining missing values were imputed using kNN from the `VIM` R package [@KowarikTempl:2016], for those metabolites that had less than `r threshold_within_nan`% of remaining missings within each cohort and `r threshold_overall_nan`% of remaining missings overall. We used `r knn_k` nearest neighbors. We natural log-transformed them to improve model fit, assessed with posterior predictive checks. To do so, replicated data were simulated with the fitted models and compared to the observed data. We used the `check_predictions` function from the `performance` R package using the default arguments [@LudeckeBen-ShacharPatil:2021]. Values of total cortisol, cortisone, and corticosterone production were expressed in nanograms per millilitre (ng/ml).

Concentrations of the non-persistent {{< acr edcs >}} were classified as quantifiable, below the {{< acr lod >}}, possible interference or out of range, and not analysed. Concentrations below the {{< acr lod >}} were singly imputed using a quantile regression approach for the imputation of left-censored missing data, as implemented in the `impute.QRILC` function from the `imputeLCMD` R package [@lazar2015imputelcmd]. Information about the lower limits of detection can be found in @HaugSakhiCequier:2018. Chemicals with more than 70% of observations below the {{< acr lod >}} were excluded from the present study. Remaining missing values were imputed similarly using kNN. Values of the chemicals were expressed in $\mu$grams per litre ($\mu$g/L).

Missing values in the clinical outcome were imputed similarly using kNN. We natural log-transformed these to improve model fit, assessed with posterior predictive checks. Values of the clinical outcome were expressed in {{< acr mss >}}.

Missing values in the covariates were imputed similarly using kNN. Categorical covariates were imputed using the `maxCat` function, which chooses the level with the most occurrences. Creatinine values were expressed in grams per litre (g/L).

### Estimation of balancing weights {#sec-weights-est}

To reduce the effect of measured confounders on the exposure-outcome association, stabilized balancing weights were estimated using the energy method available in the `WeightIt` R package [@Greifer:2023]. This method estimates weights by minimizing an energy statistic related to covariate balance [@HulingGreiferChen:2023], thus avoiding the need to specify a parametric model.
Weights below the 0.1 and above the 0.9 quantiles were trimmed. Trimming might lead to decreased covariate balance and potentially change the estimand, but can also decrease the variability of the weights.
Covariate balance was assessed using functionalities provided by the `cobalt` R package [@Greifer:2023a]. Specifically, we used *Love* plots to visualize covariate balance before and after adjusting.

### G-computation {#sec-gcomp}

We estimated {{< acr mcs >}} with the parametric g-formula, a method of standardization. The parametric g-formula involves the following steps: 1) fit a outcome model including both covariates and balancing weights; 2) create two new datasets identical to the original one but with the exposure shifted according to a user-specified intervention set by a deterministic function of the observed exposure levels; 3) use the outcome model to compute adjusted predictions in the two counterfactual datasets; 4) compute the difference between the means of the adjusted predictions in the counterfactual datasets. The causal parameter of interest was thus specified as the difference in the expected counterfactual outcomes under the shifted exposure levels $\left( \mathbb{E} \left[ Y^{d_1} \right] - \mathbb{E} \left[ Y^{d_2} \right] \right)$. In order for this parameter to be identified, the usual causal identifiability conditions (no unmeasured confounding, positivity, and consistency) are required. Since these conditions are likely not satisfied, we focused on the estimation of a statistical estimand that is as close as possible to the causal parameter of interest.

We fit the outcome model using the `glm` function and a Gaussian family with identity link from base R. The exposure variable was modeled using natural cubic splines with 3 degrees of freedom, to more flexibly capture the {{< acr adrf >}}.

To estimate the {{< acr mcs >}}, we used the `avg_comparisons` function from the `marginaleffects` R package [@Arel-Bundock:2023]. The two counterfactual datasets were obtained by setting the exposures levels to 90th percentile ($d_1$) and the 10th percentile ($d_2$), for each cohort separately. The {{< acr mcs >}} were computed using the estimated balancing weights above. Robust standard errors were computed with the `sandwich` R package, using cohort as variable indicating clustering of observations [@Zeileis:2004; @ZeileisKollGraham:2020].
For each outcome, we report the results as differences between {{< acr mcs >}}.

The R code to reproduce analyses and results is available online (https://github.com/lorenzoFabbri/paper-helixSC-neuro).

### Effect-modification analysis

We tested for possible effect-modification by sex. To do so, balancing weights were estimated separately for each level of the sex variable, and an interaction term between the exposure and sex was included in the outcome model. Similarly, the {{< acr mcs >}} were aggregated separately for each level of sex. We further tested for significance of the difference between the {{< acr mcs >}} of females and males.

# Results {#sec-res}

@tbl-pop-desc and @supptbl-pop-desc provide descriptive statistics for the outcome and covariates for the {{< acr helix >}} subcohort and for each cohort, respectively.
Of the 1,301 children of the {{< acr helix >}} subcohort, 1,297 had measurements of the non-persistent {{< acr edcs >}}. Measurements of the glucocorticosteroids were available for 1,004 children, of which 980 were matched to the {{< acr helix >}} subcohort. Measurements of both non-persistent {{< acr edcs >}} and glucocorticosteroids were available for 976 children of the subcohort. A flowchart describing the sample size for each research question is presented in @suppfig-flow-pop. The sample consisted of 55% males. The median {{< acr hitrtse >}} was 300 ms ({{< acr iqr >}}, 231-368), with lower median values for EDEN, MOBA, and INMA, corresponding to the cohorts with older children. At the time of visit, the median age of the children was 8.06 years. The children were mostly Caucasian (90%), and the largest minority were of Pakistani origin (6.2%).

Levels of unprocessed non-persistent {{< acr edcs >}}, after imputation of values below the {{< acr lod >}}, and glucocorticosteroids, are presented in @tbl-edc-desc, @tbl-met-new-desc, and @supptbl-met-desc. Supplementary Figures [-@suppfig-desc-chems] and [-@suppfig-desc-mets] provide information on the measurement classification of the {{< acr edcs >}} and glucocorticosteroids by cohort, respectively.

The effective sample sizes before and after balancing weights estimation are presented in Supplementary Tables [-@supptbl-balance-1; -@supptbl-balance-2; -@supptbl-balance-3], while basic summary statistics of the estimated balancing weights are presented in Supplementary Tables [-@supptbl-weights-1; -@supptbl-weights-2; -@supptbl-weights-3]. As expected, the median value of the weights for each exposure was close to $1.00$.

@Fig-marginal-1 presents the forest plot for the {{< acr mcs >}} of the non-persistent {{< acr edcs >}} on {{< acr hitrtse >}}. For most {{< acr edcs >}}, a cohort-specific decrease in the levels of the exposures from the 90th to the 10th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the values of {{< acr hitrtse >}} and thus lower attention. Most of the {{< acr cis first_use=true >}} included the null effect, though. Significant effects were observed for the paraben MEPA (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`). The {{< acr opp >}} DETP was negatively associated with {{< acr hitrtse >}} (`r format_res(meffects1, "detp", "HRT-SE")`).

@Fig-marginal-2 presents the forest plot for the {{< acr mcs >}} of the non-persistent {{< acr edcs >}} on total cortisone, cortisol, and corticosterone production. For most {{< acr edcs >}}, a cohort-specific decrease in the levels of the exposures from the 90th to the 10th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the total production of these metabolites. Exceptions were BUPA, which was associated with negative {{< acr mcs >}} for all three outcomes, and MiBP, which was associated with a negative {{< acr mc >}} for total cortisone production only.
The majority of the effects for the phenols and phthalate metabolites included the null. The phenol BPA showed the largest {{< acr mcs >}} across all three outcomes (cortisone production, `r format_res(meffects2, "bpa", "cortisone production")`; cortisol production, `r format_res(meffects2, "bpa", "cortisol production")`; corticosterone production, `r format_res(meffects2, "bpa", "corticosterone production")`).

@Fig-marginal-3 presents the forest plot for the {{< acr mcs >}} of the glucocorticosteroids on {{< acr hitrtse >}}. All {{< acr mcs >}} included the null, with no clear indication of directionality of the effect.

## Effect modification by sex

```{r}
#| echo: false

targets::tar_load(
  dplyr::matches("marginal_hypothesis_(\\d)SA$"),
  store = store
)
heffects1 <- tidy_res_meffects(
  df = marginal_hypothesis_1SA,
  sa_var = "e3sex",
  outcome = NULL,
  which_res = "hypothesis",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$numerical_results
heffects2 <- tidy_res_meffects(
  df = marginal_hypothesis_2SA,
  sa_var = "e3sex",
  outcome = NULL,
  which_res = "hypothesis",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$numerical_results
heffects3 <- tidy_res_meffects(
  df = marginal_hypothesis_3SA,
  sa_var = "e3sex",
  outcome = NULL,
  which_res = "hypothesis",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$numerical_results
```

Basic summary statistics of the estimated balancing weights for effect modification are presented in Supplementary Tables [-@supptbl-weights-1sa; -@supptbl-weights-2sa; -@supptbl-weights-3sa]. As expected, the median value of the weights for each exposure was close to $1.00$.

@Tbl-hypothesis-1and2 presents the results of the difference between estimates of the {{< acr mcs >}} for females and males, for the {{< acr edcs >}} on the glucocorticosteroids and {{< acr hitrtse >}}. For {{< acr hitrtse >}}, significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "HRT-SE")`) and the phthalate metabolites MEP (`r format_res(heffects1, "mep", "HRT-SE")`) and MbZP (`r format_res(heffects1, "mbzp", "HRT-SE")`). For the glucocorticosteroids, significant differences were present across all three classes of {{< acr edcs >}} and for all outcomes. The largest differences were attributable to the {{< acr opps >}} DMTP (cortisol production, `r format_res(heffects2, "dmtp", "cortisol production")`) and DETP (corticosterone production, (`r format_res(heffects2, "detp", "corticosterone production")`); cortisone production, (`r format_res(heffects2, "detp", "cortisone production")`)). The forest plots of the individual {{< acr mcs >}} are presented in Supplementary Figures [-@suppfig-marginal-1sa] and [-@suppfig-marginal-2sa].

@Tbl-hypothesis-3 presents the results of the difference between estimates of the {{< acr mcs >}} for females and males, for the glucocorticosteroids on {{< acr hitrtse >}}. Significant differences were present for cortisone production (`r format_res(heffects3, "cortisone prod.", "HRT-SE")`) and corticosterone production (`r format_res(heffects3, "corticost. prod.", "HRT-SE")`). Furthermore, for all exposures, the {{< acr mcs >}} had opposite sign (positive for males and negative for females). The forest plot of the individual {{< acr mcs >}} is presented in @suppfig-marginal-3sa.

# Discussion {#sec-discussion}

The impact of exposure to {{< acr edcs >}} on human health has attracted considerable research interest. While research in this area has mainly investigated the effects of prenatal exposure on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022], little is still known about childhood exposure. In this study, consisting of 1,297 children from 6 European birth cohorts, we observed that short-term childhood exposure to certain non-persistent {{< acr edcs >}} was associated with attentional function (MEPA, MEHP, oh-MiNP, and oxo-MiNP), and with total production of cortisol, cortisone, and corticosterone (DEP, DMP, DMTP, BPA, ETPA, MEPA, MEHP, oh-MiNP, and oxo-MiNP). Increased production of these glucocorticosteroids did not seem to affect attentional function. Some of these effects differed for females and males, including significant differences for the effects of increased production of cortisone and corticosterone on {{< acr hitrtse >}}. Specifically, an increased production of these glucocorticosteroids was associated with lower values of {{< acr hitrtse >}} for females, and higher values for males. Taken together, these results suggest that these non-persistent {{< acr edcs >}} might be responsible for perturbations of the {{< acr hpa >}} axis' homeostasis, and that higher levels of these glucocorticosteroids might interfere with different functions of attention in a sex-specific manner.

===
To the best of our knowledge, no other study has investigated the effects of childhood exposure to multiple classes of non-persistent {{< acr edcs >}} in relation to attentional function.

More generally, the literature on non-persistent {{< acr edcs >}} and neurodevelopment in children has mostly focused on {{< acr opps >}} [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @CartierWarembourgLeManer-Idrissi:2016; @YuDuChiou:2016], phthalate metabolites [@HuangChenSu:2015; @HuangTsaiChen:2017; @KimHongShin:2017; @LiPapandonatosCalafat:2019; @BalalianWhyattLiu:2019; @ShoaffCoullWeuve:2020; @JankowskaPolanskaHanke:2019; @VilmandBeckBilenberg:2023], and {{< acr bpa >}} [@TewarAuingerBraun:2016; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019].

González-Alzaga et al. and Cartier et al. evaluated cross-sectional associations between {{< acr dap >}} metabolites and subtests of the Wechsler Intelligence Scale for Children [@wechsler1949wechsler] in European children with ages between 6 and 11 years. Higher levels of {{< acr dap >}} metabolites (DMP, DMTP, DMDTP, DEP, DETP, and DEDTP) were associated with lower scores of {{< acr iq >}} and verbal comprehension, especially in boys [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015], while higher levels of diethylphosphate metabolites (DEP, DETP, DEDTP) were associated with lower working memory scores [@CartierWarembourgLeManer-Idrissi:2016].
There is also preliminary evidence of a possible association between exposure to certain {{< acr opps >}} and {{< acr adhd >}} in children [@BouchardBellingerWright:2010; @YuDuChiou:2016]. Specifically, Bouchard et al. found evidence of a cross-sectional association between dimethyl alkylphosphate metabolites (DMP, DMTP, and DMDTP) and {{< acr adhd >}} in children aged 8 to 15 years from {{< acr nhanes >}}, while Yu et al. found a dose-response relationship between DMP and {{< acr adhd >}} in Taiwanese children aged 4 to 15 years.

Preliminary evidence is also available for several phthalate metabolites in relation to cognitive development in childhood. Higher levels of di(2-ethylhexyl) phthalate metabolites (including MEHP, MEHHP, and MEOHP) were associated with lower intelligence scores in children aged 2 to 12 years [@HuangChenSu:2015], lower scores of {{< acr iq >}} and verbal {{< acr iq >}}, more omission errors (a measure of inattention), and higher scores of response time variability (a measure of sustained attention) in 6-year old Korean children [@KimHongShin:2017], and poorer fine motor skills in preadolescent boys [@BalalianWhyattLiu:2019].

Few studies have looked into different classes of non-persistent {{< acr edcs >}} [@ShoaffCoullWeuve:2020 in relation with the Conners Attention Deficit Scale and the Behavior Assessment System for Children; @OhKimKannan:2023 in relation with {{< acr adhd >}} symptoms]. Overall, and consistent with our results, these studies seem to provide further evidence of the adverse effects of several non-persistent {{< acr edcs >}} on neurodevelopment (too broad) in children. While not all these studies have investigated effect modification by sex, it seems that these adverse effects tend to be stronger in males [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @YuDuChiou:2016; @ShoaffCoullWeuve:2020], although this difference might depend on the timing of exposure [@].
===

Our results are consistent with prior epidemiological research that associated exposure to certain non-persistent {{< acr edcs >}} with higher levels of cortisol [@KimLeeMoon:2018; @SunLiJin:2018; @SearsLiuLanphear:2023]. There are some differences, though. First, these studies only focus on phthalates, either as individual metabolites or as mixtures. Second, exposure assessment in @SunLiJin:2018 and @KimLeeMoon:2018 was performed during gestation or the first 15 months of life, respectively. Finally, the glucocorticosteroids were measured in cord blood [@SunLiJin:2018] and hair [@SearsLiuLanphear:2023]. Contrary to these studies, we did find effect modification by sex. We are not aware of other epidemiological studies investigating phthalates metabolites, phenols, and {{< acr opps >}}, in relation to urinary glucocorticosteroids in childhood. Nonetheless, previous toxicological studies provide evidence for the inhibition by phthalates of human 11$\beta$-hydroxysteroid dehydrogenase 2 (11$\beta$-HSD2) activity, responsible for the conversion of active cortisol into inactive cortisone [@ZhaoChuHuang:2010; @MaLianDong:2011]. There is also *in silico* evidence suggesting that {{< acr bpa >}}, a phenol, and Triazophos (TAP), a organophosphorus insecticide, can bind to the human glucocorticoid receptor [@PrasanthDivyaSadasivan:2010; @YangLiRen:2019].

We are also not aware of prior epidemiological studies specifically investigating the effects of elevated levels of glucocorticosteroids in relation to attentional function, although there is evidence that under- or over-production of glucocorticosteroids interfere with the normal development of the brain [@LupienMcEwenGunnar:2009]. While we did find sex-specific evidence of an effect, their clinical relevance is questionable.

Our findings should be interpreted in light of the following limitations and strengths. Limitations include the cross-sectional design of the present study. Importantly, the non-persistent {{< acr edcs >}} were measured in a pool of night and morning urine samples before the clinical visit, to represent exposure over the previous day, whereas the glucocorticosteroids were measured in the night urine sample. Although we included a wide range of confounders there is the possibility, as with other observational studies, of residual confounding, which might lead to a bias away from the null. Some of the confounders indicated in the adjustment sets had to be removed due to large fractions of missing values. There is further the possibility of misspecification of the outcome model, although we included a spline of the exposure to relax some of the linearity assumptions. The use of more data-adaptive learners was excluded due to the relatively small sample size. We finally acknowledge the possibility that some of chemicals might not act independently (mixture effect). Further research is thus warranted.

Strengths of the present study include the use of pooled urine samples for chemical assessment to obtain more representative long-term exposures, since it is known that these specific {{< acr edcs >}} have very short half-lives [@PerrierGiorgis-AllemandSlama:2016; @CasasBasaganaSakhi:2018]. We decided to model both the *treatment* mechanisms, for the estimation of balancing weights, and the outcomes, with traditional covariates adjustment, to try to obtain *doubly robust* effect estimates. Finally, we decided not to interpret our results by focusing on the estimated coefficients of possibly misspecified regression models, but by making use of the g-computation procedure and estimate {{< acr mcs >}}.

In conclusion, in a study of 1,297 children from 6 European birth cohorts, we observed that (i) exposure to non-persistent {{< acr edcs >}} in childhood might have short-term effects on {{< acr hitrtse >}} in childhood, (ii) exposure to non-persistent {{< acr edcs >}} in childhood might disrupt the {{< acr hpa >}} axis in childhood, and (iii) disruption of the {{< acr hpa >}} axis in childhood might have short-term, sex-specific effects on {{< acr hitrtse >}}. Future studies should investigate how glucocorticosteroids might mediate the adverse effects of exposure to non-persistent {{< acr edcs >}} on childhood neurodevelopment (too broad) in larger populations.

{{< pagebreak >}}

# References {.unnumbered}

::: {#refs}
:::

{{< pagebreak >}}

{{< include _tables.qmd >}}

{{< pagebreak >}}

{{< include _figures.qmd >}}

{{< pagebreak >}}

{{< include _si.qmd >}}
